---
title: "*Spatial Statistics Lab 8*"
author: '**Onyedikachi J Okeke**'
date: "`r Sys.Date()`"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### **0.0 To Load the library**
```{r,comment="",echo=TRUE,eval=FALSE}
library(sp)
library(spatstat)
library(sf)
library(spatstat.geom)
library(ctv)
library(terra)
library(spdep)
library(rgdal)
library(rgdal)
library(terra)
library(RColorBrewer)
library(classInt)
library(epitools)
library(DCluster)
library(lmtest)
library(car)
library(spatialreg)
library(spdep)
library(classInt)
library(spgwr)
library(ggplot2)

```
##### Result
```{r echo=FALSE, results='show', fig.show='hold', eval=TRUE, warning=FALSE, fig.width=6, fig.height=4}

library(sp)
library(spatstat)
library(sf)
library(spatstat.geom)
library(ctv)
library(terra)
library(spdep)
library(rgdal)
library(rgdal)
library(terra)
library(RColorBrewer)
library(classInt)
library(epitools)
library(DCluster)
library(lmtest)
library(car)
library(spatialreg)
library(spdep)
library(classInt)
library(spgwr)
library(ggplot2)

```

### **0.1 Loading Boston data**
```{r,comment="",echo=TRUE,eval=FALSE}

Boston <- readOGR(system.file("shapes/boston_tracts.shp", package="spData")[1])
class(Boston)
plot(Boston)

```
##### Result
```{r echo=FALSE, results='show', fig.show='hold', eval=TRUE, warning=FALSE, fig.width=6, fig.height=4}

Boston <- readOGR(system.file("shapes/boston_tracts.shp", package="spData")[1])
class(Boston)
plot(Boston)

```

## **Part One**

### **0.1 Row-standardized data and OLS for Housing**
```{r,comment="",echo=TRUE,eval=FALSE}

# creates a row-standardized "listw" object (default style = "W")
Boston_queen_nb <- poly2nb(Boston, queen=TRUE)
Boston_w <- nb2listw(Boston_queen_nb)

# creates a non row-standardized "listw" object
Boston_b <- nb2listw(Boston_queen_nb,style="B")

# displays the "data.frame" attribute of the shape file

boston_data <- attr(Boston, "data")
# Extracting the first 10 rows
first_5 <- head(boston_data, 5)
# Extracting the last 10 rows
last_5 <- tail(boston_data, 5)
# Combining the first 10 and last 10 rows
combined_rows <- rbind(first_5, last_5)

# Display the combined rows
print(combined_rows)

# using log of median housing value instead of housing value
Boston$LOGMEDV<-log(Boston$CMEDV)

# OLS for housing value
Boston_OLS <- lm(LOGMEDV~CRIM + CHAS + NOX + RM + AGE + DIS + B + LSTAT, data=Boston)
summary(Boston_OLS)
# saving residuals and fitted values
Boston$olsresid<-residuals(Boston_OLS)
Boston$ols_fitted <- fitted(Boston_OLS)
# Moran test for residuals
lm.morantest(Boston_OLS, Boston_w)

```
##### Result
```{r echo=FALSE, results='show', fig.show='hold', eval=TRUE, warning=FALSE, fig.width=6, fig.height=4}

# creates a row-standardized "listw" object (default style = "W")
Boston_queen_nb <- poly2nb(Boston, queen=TRUE)
Boston_w <- nb2listw(Boston_queen_nb)

# creates a non row-standardized "listw" object
Boston_b <- nb2listw(Boston_queen_nb,style="B")

# displays the "data.frame" attribute of the shape file

boston_data <- attr(Boston, "data")
# Extracting the first 10 rows
first_10 <- head(boston_data, 10)
# Extracting the last 10 rows
last_10 <- tail(boston_data, 10)
# Combining the first 10 and last 10 rows
combined_rows <- rbind(first_10, last_10)

# Display the combined rows
print(combined_rows)

# using log of median housing value instead of housing value
Boston$LOGMEDV<-log(Boston$CMEDV)

# OLS for housing value
Boston_OLS <- lm(LOGMEDV~CRIM + CHAS + NOX + RM + AGE + DIS + B + LSTAT, data=Boston)
summary(Boston_OLS)
# saving residuals and fitted values
Boston$olsresid<-residuals(Boston_OLS)
Boston$ols_fitted <- fitted(Boston_OLS)
# Moran test for residuals
lm.morantest(Boston_OLS, Boston_w)

```

Q1. Describe the OLS regression model results and their significance.

### **Lagrange multiplier tests for residuals**
```{r,comment="",echo=TRUE,eval=FALSE}

lm.LMtests(Boston_OLS, Boston_w, test="all")
#Breusch Pagan test for heteroskedasticity
bptest(Boston_OLS)

```
##### Result
```{r echo=FALSE, results='show', fig.show='hold', eval=TRUE, warning=FALSE, fig.width=6, fig.height=4}

lm.LMtests(Boston_OLS, Boston_w, test="all")
#Breusch Pagan test for heteroskedasticity
bptest(Boston_OLS)

```

Q2. Describe the diagnostics test results for spatial autocorrelation.

### **Spatial lag Regression**

```{r,comment="",echo=TRUE,eval=FALSE}

Boston_spatial_lag <- lagsarlm(LOGMEDV ~ CRIM + CHAS + NOX + RM + AGE + DIS + B + LSTAT, data=Boston, Boston_w)
summary(Boston_spatial_lag)
# saving residuals and fitted values
Boston$lagresid<-residuals(Boston_spatial_lag)
Boston$lag_fitted <- fitted(Boston_spatial_lag)

```
##### Result
```{r echo=FALSE, results='show', fig.show='hold', eval=TRUE, warning=FALSE, fig.width=6, fig.height=4}

Boston_spatial_lag <- lagsarlm(LOGMEDV ~ CRIM + CHAS + NOX + RM + AGE + DIS + B + LSTAT, data=Boston, Boston_w)
summary(Boston_spatial_lag)
# saving residuals and fitted values
Boston$lagresid<-residuals(Boston_spatial_lag)
Boston$lag_fitted <- fitted(Boston_spatial_lag)

```

Q3. Describe the spatial lag regression model results and its significance.

### **Error regression** 
```{r,comment="",echo=TRUE,eval=FALSE}

Boston_spatial_error <- errorsarlm(LOGMEDV ~ CRIM + CHAS + NOX + RM + AGE + DIS + B + LSTAT, data=Boston, Boston_w)
summary(Boston_spatial_error)
# saving residuals and fitted values
Boston$error_resid<-residuals(Boston_spatial_error)
Boston$error_fitted <- fitted(Boston_spatial_error)

```
##### Result
```{r echo=FALSE, results='show', fig.show='hold', eval=TRUE, warning=FALSE, fig.width=6, fig.height=4}

Boston_spatial_error <- errorsarlm(LOGMEDV ~ CRIM + CHAS + NOX + RM + AGE + DIS + B + LSTAT, data=Boston, Boston_w)
summary(Boston_spatial_error)
# saving residuals and fitted values
Boston$error_resid<-residuals(Boston_spatial_error)
Boston$error_fitted <- fitted(Boston_spatial_error)

```

Q4. Describe the spatial error regression model results and its significance.

## **Part Two**
### **create a data frame** 
```{r,comment="",echo=TRUE,eval=FALSE}

# create a data frame
boston.df <- data.frame(Boston)
names(boston.df) # gives us the column names
colnames(boston.df) # alternative call; same thing

attach(boston.df) #Now you can call the variable name without using “$”

LOGCMEDV <- log(CMEDV)

# we create this location using the coordinates of the centroids:
XYgridtable <- cbind(LON, LAT)
#using gwr.sel to find optimal parameters
adaptive.bw<-gwr.sel(LOGCMEDV~CRIM+CHAS+NOX+RM+AGE+DIS+B+LSTAT,adapt=TRUE, method="cv", coords=XYgridtable)
###gwr.Gauss() default

#number of observations (polygon)
dim(boston.df)[1]
# how many points does each window take
print(dim(boston.df)[1]*adaptive.bw )
#run gwr using the optimal parameter found above
bos.gwr<-gwr(LOGCMEDV ~ CRIM + CHAS + NOX + RM + AGE + DIS + B + LSTAT, adapt = adaptive.bw, gweight = gwr.Gauss, hatmatrix = TRUE, coords = XYgridtable)
print(bos.gwr)
#Plot local R2
classes_fx <- classIntervals(bos.gwr$SDF$localR2, n=5, style="quantile")
cols <- findColours(classes_fx,pal)
pal <- brewer.pal(5, "YlOrRd")
plot(Boston,col=cols, border="grey")
legend(x="topleft", cex=0.65, fill=attr(cols,"palette"), bty="n", legend=names(attr(cols, "table")), title="Local R^2",ncol=1)
#Plot Residual
res <- bos.gwr$SDF$gwr.e
classes_fx <- classIntervals(res, n=5, style="quantile")
pal <- brewer.pal(5, "YlOrRd")
cols <- findColours(classes_fx,pal)
plot(Boston,col=cols, border="transparent")
legend(x="topleft", cex=0.65, fill=attr(cols, "palette"), bty="n", legend=names(attr(cols, "table")), title="Residual", ncol=1)

```
##### Result
```{r echo=FALSE, results='show', fig.show='hold', eval=TRUE, warning=FALSE, fig.width=6, fig.height=4}

# create a data frame
boston.df <- data.frame(Boston)
names(boston.df) # gives us the column names
colnames(boston.df) # alternative call; same thing
attach(boston.df) #Now you can call the variable name without using “$”
LOGCMEDV <- log(CMEDV)

# we create this location using the coordinates of the centroids:
XYgridtable <- cbind(LON, LAT)
#using gwr.sel to find optimal parameters
adaptive.bw<-gwr.sel(LOGCMEDV~CRIM+CHAS+NOX+RM+AGE+DIS+B+LSTAT,adapt=TRUE, method="cv", coords=XYgridtable)
###gwr.Gauss() default

#number of observations (polygon)
dim(boston.df)[1]
# how many points does each window take
print(dim(boston.df)[1]*adaptive.bw )
#run gwr using the optimal parameter found above
bos.gwr<-gwr(LOGCMEDV ~ CRIM + CHAS + NOX + RM + AGE + DIS + B + LSTAT, adapt = adaptive.bw, gweight = gwr.Gauss, hatmatrix = TRUE, coords = XYgridtable)

print(bos.gwr)
#Plot local R2
classes_fx <- classIntervals(bos.gwr$SDF$localR2, n=5, style="quantile")

pal <- brewer.pal(5, "YlOrRd")

cols <- findColours(classes_fx,pal)

plot(Boston,col=cols, border="grey")
legend(x="topleft", cex=0.65, fill=attr(cols,"palette"), bty="n", legend=names(attr(cols, "table")), title="Local R^2",ncol=1)
#Plot Residual
res <- bos.gwr$SDF$gwr.e
classes_fx <- classIntervals(res, n=5, style="quantile")
pal <- brewer.pal(5, "YlOrRd")
cols <- findColours(classes_fx,pal)
plot(Boston,col=cols, border="transparent")
legend(x="topleft", cex=0.65, fill=attr(cols, "palette"), bty="n", legend=names(attr(cols, "table")), title="Residual", ncol=1)

```

Q5. Take screenshots of the plots and describe the pattern of local R2 and residual.

### **Spatial lag Regression**

```{r,comment="",echo=TRUE,eval=FALSE}

###plot local coefficient (CRIM)
coef <- bos.gwr$SDF$CRIM
classes_fx <- classIntervals(coef, n=5, style="quantile")
cols <- findColours(classes_fx,pal)
plot(Boston,col=cols, border="transparent")
legend(x="topleft", cex=0.75, fill=attr(cols, "palette"), bty="n", legend=names(attr(cols, "table")), title = "Local Coefficient Estimates (CRIM)", ncol=1)

###plot local coefficient (NOX)
coef <- bos.gwr$SDF$NOX
classes_fx2 <- classIntervals(coef, n=5, style="quantile")
cols <- findColours(classes_fx2,pal)
plot(Boston,col=cols, border="transparent")
legend(x="topleft", cex=0.7, fill=attr(cols, "palette"), bty="n", legend=names(attr(cols, "table")), title="Local Coefficient Estimates (NOX)", ncol = 1)

#Plot RM, AGE, DIS, B, LSTAT

###plot local coefficient (RM)
coef <- bos.gwr$SDF$RM
classes_fx4 <- classIntervals(coef, n=5, style="quantile")
cols <- findColours(classes_fx4,pal)
plot(Boston,col=cols, border="transparent")
legend(x="topleft", cex=0.7, fill=attr(cols, "palette"), bty="n", legend=names(attr(cols, "table")), title="Local Coefficient Estimates (RM)", ncol = 1)

###plot local coefficient (AGE)
coef <- bos.gwr$SDF$AGE
classes_fx5 <- classIntervals(coef, n=5, style="quantile")
cols <- findColours(classes_fx5,pal)
plot(Boston,col=cols, border="transparent")
legend(x="topleft", cex=0.7, fill=attr(cols, "palette"), bty="n", legend=names(attr(cols, "table")), title="Local Coefficient Estimates (AGE)", ncol = 1)


###plot local coefficient (DIS)
coef <- bos.gwr$SDF$DIS
classes_fx6 <- classIntervals(coef, n=5, style="quantile")
cols <- findColours(classes_fx6,pal)
plot(Boston,col=cols, border="transparent")
legend(x="topleft", cex=0.7, fill=attr(cols, "palette"), bty="n", legend=names(attr(cols, "table")), title="Local Coefficient Estimates (DIS)", ncol = 1)


###plot local coefficient (B)
coef <- bos.gwr$SDF$B
classes_fx7 <- classIntervals(coef, n=5, style="quantile")
cols <- findColours(classes_fx7,pal)
plot(Boston,col=cols, border="transparent")
legend(x="topleft", cex=0.7, fill=attr(cols, "palette"), bty="n", legend=names(attr(cols, "table")), title="Local Coefficient Estimates (B)", ncol = 1)

```
##### Result
```{r echo=FALSE, results='show', fig.show='hold', eval=TRUE, warning=FALSE, fig.width=8, fig.height=5}

###plot local coefficient (CRIM)
coef <- bos.gwr$SDF$CRIM
classes_fx <- classIntervals(coef, n=5, style="quantile")
cols <- findColours(classes_fx,pal)
plot(Boston,col=cols, border="transparent")
legend(x="topleft", cex=0.75, fill=attr(cols, "palette"), bty="n", legend=names(attr(cols, "table")), title = "Local Coefficient Estimates (CRIM)", ncol=1)

###plot local coefficient (NOX)
coef <- bos.gwr$SDF$NOX
classes_fx2 <- classIntervals(coef, n=5, style="quantile")
cols <- findColours(classes_fx2,pal)
plot(Boston,col=cols, border="transparent")
legend(x="topleft", cex=0.7, fill=attr(cols, "palette"), bty="n", legend=names(attr(cols, "table")), title="Local Coefficient Estimates (NOX)", ncol = 1)

#Plot RM, AGE, DIS, B, LSTAT

###plot local coefficient (RM)
coef <- bos.gwr$SDF$RM
classes_fx4 <- classIntervals(coef, n=5, style="quantile")
cols <- findColours(classes_fx4,pal)
plot(Boston,col=cols, border="transparent")
legend(x="topleft", cex=0.7, fill=attr(cols, "palette"), bty="n", legend=names(attr(cols, "table")), title="Local Coefficient Estimates (RM)", ncol = 1)

###plot local coefficient (AGE)
coef <- bos.gwr$SDF$AGE
classes_fx5 <- classIntervals(coef, n=5, style="quantile")
cols <- findColours(classes_fx5,pal)
plot(Boston,col=cols, border="transparent")
legend(x="topleft", cex=0.7, fill=attr(cols, "palette"), bty="n", legend=names(attr(cols, "table")), title="Local Coefficient Estimates (AGE)", ncol = 1)


###plot local coefficient (DIS)
coef <- bos.gwr$SDF$DIS
classes_fx6 <- classIntervals(coef, n=5, style="quantile")
cols <- findColours(classes_fx6,pal)
plot(Boston,col=cols, border="transparent")
legend(x="topleft", cex=0.7, fill=attr(cols, "palette"), bty="n", legend=names(attr(cols, "table")), title="Local Coefficient Estimates (DIS)", ncol = 1)


###plot local coefficient (B)
coef <- bos.gwr$SDF$B
classes_fx7 <- classIntervals(coef, n=5, style="quantile")
cols <- findColours(classes_fx7,pal)
plot(Boston,col=cols, border="transparent")
legend(x="topleft", cex=0.7, fill=attr(cols, "palette"), bty="n", legend=names(attr(cols, "table")), title="Local Coefficient Estimates (B)", ncol = 1)

```

### **plot t values (CRIME)** 
```{r,comment="",echo=TRUE,eval=FALSE}

#plot t values (CRIME)
bos.gwr$SDF$CRIM_t <- bos.gwr$SDF$CRIM/bos.gwr$SDF$CRIM_se
tCRIM<-bos.gwr$SDF$CRIM_t
display.brewer.pal(3, "RdBu")
colors <- brewer.pal(3, "RdBu") # Stores colors in object color
color.category.reg <- findInterval(tCRIM,
c(min(tCRIM)-.0001, -1.96, -1.64,
max(tCRIM)+.0001),all.inside=TRUE)
#classes_fx <- classIntervals(tCRIM, n=5, style="quantile")
#cols <- findColours(classes_fx,pal)
plot(Boston,col=colors[color.category.reg], border="transparent")
labels <- c("<-1.96", " -1.96 to -1.64", "Not significant")
legend("bottom", legend=labels, fill=colors, cex=0.8,
y.intersp = 0.99, bty="n")

```
##### Result
```{r echo=FALSE, results='show', fig.show='hold', eval=TRUE, warning=FALSE, fig.width=8, fig.height=5}

library(ggplot2)

#plot t values (CRIME)
bos.gwr$SDF$CRIM_t <- bos.gwr$SDF$CRIM/bos.gwr$SDF$CRIM_se
tCRIM<-bos.gwr$SDF$CRIM_t
display.brewer.pal(3, "RdBu")
colors <- brewer.pal(3, "RdBu") # Stores colors in object color
color.category.reg <- findInterval(tCRIM,
c(min(tCRIM)-.0001, -1.96, -1.64,
max(tCRIM)+.0001),all.inside=TRUE)

#classes_fx <- classIntervals(tCRIM, n=5, style="quantile")
#cols <- findColours(classes_fx,pal)
plot(Boston,col=colors[color.category.reg], border="transparent")
labels <- c("<-1.96", " -1.96 to -1.64", "Not significant")
legend("bottom", legend=labels, fill=colors, cex=0.8,
y.intersp = 0.99, bty="n")


# Plot the t-values using spplot
spplot(bos.gwr$SDF, "CRIM_t", main="t-values for CRIM Coefficients",
       col.regions=colorRampPalette(c("blue", "white", "red"))(100))

# Generate summary for the GWR model
bos.gwr.summary <- summary(bos.gwr, verbose = TRUE)  # the verbose = TRUE might provide more info

# Look at the structure of the summary to find coefficients and standard errors
str(bos.gwr.summary)


```





Q7. Take a screen shot of the plot and describe the pattern. (10 points)
Q8. Describe the GWR model results and their significance. How does it compare with the regression results in part 1? (10 points)
Q9. How could you use these results as a policymaker who is interested in environmental or social issues (pollution or affordable housing)?